Sperm should evolve to make female meiosis fair.

### 2014-07-28 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
fixed couple of citations

Because female meiotic drive is likely a common force with predictably negative effects on organismal fitness,  
	and because sperm have ample opportunity to influence female drive,  
	we develop population genetic models to address the the expected influence of sperm on female drive.  
We first focus on models in which `self-promoting' alleles in
  sperm facilitate drive of like alleles during gametogenesis in heterozygous females. 
These models show that such sperm-acting alleles 
	have more difficulty invading a population than do traditional meiotic drivers, 
	 and under most circumstances, cannot be maintained as a balanced polymorphism.
Because self-promoting drivers are unlikely to create a sustained genomic conflict, 
	female meiosis will have little incentive to evolve resistance to them.
We then examine models in which a novel sperm-acting allele modifies the efficacy of a polymorphic meiotic driver. 
Such models universally reveal that sperm-acting drive modifiers are favored only if they suppress drive. 
These results highlight the fact that the interests of sperm and
maternal  genomes' are often aligned, 
as both are invested in the fate of the resultant zygote \citep[as was speculated for the \emph{In} locus ][]{Pomiankowski1993}.
Thus, there is little selective benefit to females in preventing sperm to influence female meioses, 
	and in fact, females eschewing this delay would potentially lose sperm assistance in the suppression of meiotic drivers.  
Given the wide-spread requirement of fertilization for the completion
	of  female meiosis, various features of the interaction between sperm 
	and egg may function to enforce an equitable transfer of genetic material.

@@

Sexual reproduction is a high-stakes event that determines what gets transmitted to the next generation.  
As a consequence of this intense competition, alleles that gain a transmission advantage during reproduction 
	can succeed evolutionarily even if they lower organismal fitness. 
This generates numerous conflicts  \citep{Burt2006} including sexual conflicts between mates \citep{Arnqvist2005}, 
	and conflicts between alleles that are over-transmitted in meiosis and the organisms they injure while doing so. 
Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \citep{Rice2013}. \newline

@@

The theory presented herein provides a strong prediction -- 
	for organisms in which female  meiosis is not completed until after fertilization, 
	sperm will act to suppress female drive at the stage at which they can influence meiosis. 
%Therefore, for example, when fertilized by drying sperm, we expect meiosis to be fairer when  females are fertilized by 
Contrary to expectations of a naive verbal ``green-beard'' model, 
	our results are consistent with the observation that female meiosis 
	is fairer when fertilized by sperm bearing the drive allele in 
	two of the best described cases of female meiotic drive in 
	animals \citep[the \emph{Om} and \emph{In} loci in mice][]{Agulnik1993,Wu2005}.  
Given the benefit to sperm of hampering female drive, 
	drive-suppressing sperm are often likely to be fixed within a species, 
	making the hypothesis of sperm-acting drive-suppression difficult to test from intra-population crosses. 
However, crosses between populations or species are likely to provide critical tests of our theory -- 
	specifically we predict that female meiosis will be less fair when a species (or population) is 
	fertilized by heterospecific sperm because such sperm have not evolved to counter novel females meiotic drivers. 
We further predict that segregation in F1 females backcrossed to parental species will likely be biased, 
	with a deficit of transmission of the paternal species allele from the F1  female.   
Tthese predictions follow straightforwardly from the
	theory presented above; however, we caution that tests of meiotic drive, 
	and especially sperm-dependent meiotic drive require a high standard
	of evidence  to exclude plausible alternative hypotheses such as genotypic 
	inviability including epistatic maternal X zygotic lethality \citep[e.g. ][]{Sawamura:1993aa}.  \newline

>>>>

Because female meiotic drive is likely a common force with predictably negative effects on organismal fitness,  
	and because sperm have ample opportunity to influence female drive,  
	we develop population genetic models to address the the expected influence of sperm on female drive.  
We first focus on models in which `self-promoting' alleles in
  sperm facilitate drive of like alleles during gametogenesis in heterozygous females. 
These models show that such sperm-acting alleles 
	have more difficulty invading a population than do traditional meiotic drivers, 
	 and under most circumstances, cannot be maintained as a balanced polymorphism.
Because self-promoting drivers are unlikely to create a sustained genomic conflict, 
	female meiosis will have little incentive to evolve resistance to them.
We then examine models in which a novel sperm-acting allele modifies the efficacy of a polymorphic meiotic driver. 
Such models universally reveal that sperm-acting drive modifiers are favored only if they suppress drive. 
These results highlight the fact that the interests of sperm and
maternal  genomes' are often aligned, 
as both are invested in the fate of the resultant zygote \citep[as was speculated for the \emph{In} locus, ][]{Pomiankowski1993}.
Thus, there is little selective benefit to females in preventing sperm to influence female meioses, 
	and in fact, females eschewing this delay would potentially lose sperm assistance in the suppression of meiotic drivers.  
Given the wide-spread requirement of fertilization for the completion
	of  female meiosis, various features of the interaction between sperm 
	and egg may function to enforce an equitable transfer of genetic material.

@@

Sexual reproduction is a high-stakes event that determines what gets transmitted to the next generation.  
As a consequence of this intense competition, alleles that gain a transmission advantage during reproduction 
	can succeed evolutionarily even if they lower organismal fitness. 
This generates numerous conflicts including sexual conflicts between mates \citep{Arnqvist2005}, 
	and conflicts between alleles that are over-transmitted in
        meiosis and the organisms they injure while doing so  \citep{Burt2006}. 
Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \citep{Rice2013}. \newline

@@

The theory presented herein provides a strong prediction -- 
	for organisms in which female  meiosis is not completed until after fertilization, 
	sperm will act to suppress female drive at the stage at which they can influence meiosis. 
%Therefore, for example, when fertilized by drying sperm, we expect meiosis to be fairer when  females are fertilized by 
Contrary to expectations of a naive verbal ``green-beard'' model, 
	our results are consistent with the observation that female meiosis 
	is fairer when fertilized by sperm bearing the drive allele in 
	two of the best described cases of female meiotic drive in 
	animals \citep[the \emph{Om} and \emph{In} loci in mice, ][]{Agulnik1993,Wu2005}.  
Given the benefit to sperm of hampering female drive, 
	drive-suppressing sperm are often likely to be fixed within a species, 
	making the hypothesis of sperm-acting drive-suppression difficult to test from intra-population crosses. 
However, crosses between populations or species are likely to provide critical tests of our theory -- 
	specifically we predict that female meiosis will be less fair when a species (or population) is 
	fertilized by heterospecific sperm because such sperm have not evolved to counter novel females meiotic drivers. 
We further predict that segregation in F1 females backcrossed to parental species will likely be biased, 
	with a deficit of transmission of the paternal species allele from the F1  female.   
Tthese predictions follow straightforwardly from the
	theory presented above; however, we caution that tests of meiotic drive, 
	and especially sperm-dependent meiotic drive require a high standard
	of evidence  to exclude plausible alternative hypotheses such as genotypic 
	inviability including epistatic maternal X zygotic lethality \citep[e.g. ][]{Sawamura:1993aa}.  \newline
